### 2023-08-26 chapters/paper_retrospective.tex
Updates
Write more on the introduction

\chapter[a retrospective study on machine learning-assisted stroke recognition for medical helpline calls]{A Retrospective Study on Machine Learning-Assisted Stroke Recognition for Medical Helpline Calls}
\label{chp:paper-retrospective}
% \ifthenelse{\equal{\skippapers}{true}}{}{

@@

\section{Introduction}
Stroke is a leading cause of disability and death worldwide \cite{cite1,cite2,cite3}. Effective treatment is time- sensitive, and an optimal outcome is more likely when treatment is administered within the first four and a half hours from stroke onset \cite{cite4,cite5}. The gateway to ambulance transport and hospital admittance is through prehospital telehealth services, including emergency medical call centres, nurse advice call lines, and out-of-hours health services. In the pre-hospital setting, the use of Mobile Stroke Units has made it possible to deliver advanced treatment faster \cite{cite6,cite7}. As the Mobile Stroke Unit is only dispatched to patients with a suspected stroke, the impact of Mobile Stroke Units is directly influenced by accurate call-taker recognition of stroke.6,7 Call-takers who can rapidly and accurately recognize stroke are therefore crucial in facilitating prompt care in both pre-hospital and in-hospital settings.

Despite initiatives to improve stroke recognition \cite{cite8,cite9}, approximately half of all patients with stroke do not receive the correct triage for their condition from call-takers \cite{cite10,cite11,cite12}. Most initiatives aim to improve stroke recognition by call-takers via introducing more specific assessment tools \cite{cite8,cite9} or providing specialized training \cite{cite13}. Recent advances in machine learning technology might be applied to improve stroke recognition without requiring changes to the triaging approach, and machine learning aided identification of stroke has been suggested as a means of improving Mobile Stroke Unit effectiveness \cite{cite7}. Real-time feedback from a machine learning model can improve the recognition of out-of-hospital cardiac arrest \cite{cite14,cite15}. Therefore, this study aimed to develop and assess the potential of machine learning in improving prehospital stroke recognition during medical helpline calls.

@@

This study has some limitations. First, the mapping of call recordings to electronic records was incomplete due to technical limitations in the CAD registry, which limited the number of calls available to us. Of note, there was no obvious pattern of bias related to the unmapped calls, and we included all calls with matching audio files, regardless of dispatcher performance. The results could potentially be improved if more calls were available for analysis. Second, calls without a call-taker indicated diagnostic category were not included in the validation and test data because the call-taker's performance could not be evaluated. Moreover, in exploratory analyses, the model performed worse on these calls, which might be attributed to differences in population characteristics (\cref{tab_retrospective:table3}). Finally, the ground truth stroke labelling relied on the patient-reported time of onset being exact; however, estimating the accuracy of the timestamps in DanStroke was impossible..

@@

In conclusion, using the largest collection of audio calls from patients with stroke to date, we developed a machine learning framework that significantly outperformed human call-takers in stroke recognition in medical helpline calls. The frameworkcan assist human call-takers during medical helpline calls. Ideally, this would enable a higher recognition of patients with stroke in the prehospital setting, benefiting both patient outcomes and health service resource allocation.

@@

JW, JDH and LB share lead authorship and contributed equally to the original drafting of the paper, review and editing, data curation, formal analysis, methodology, validation, and conceptualisation. JDH and LB contributed to software development and funding acquisition. SNFB contributed to the conceptualisation, data curation, funding acquisition, supervision, and writing - review and editing. LM contributed to conceptualisation, funding acquisition, project administration, supervision, and writing - review and editing. MS contributed to supervision and writing - review and editing. HC contributed to the methodology, supervision, and writing - review and editing. CK contributed to the conceptualisation, funding acquisition, methodology, project administration, supervision, and writing - review and editing. HCC contributed to the conceptualisation, data curation, funding acquisition, methodology, project administration, supervision, and writing - review and editing.

HCC and CK share last authorship.

The three lead authors directly accessed and verified the underlying data reported in this manuscript. JW's affiliations are exclusively academic.

@@

% \todo[inline]{Insert Table 3}
\begin{table}[t]
    \centering
    \caption[Words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.]{English translation of words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.}
    \label{tab_retrospective:table3}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|lr|lr}
        \toprule
                    & \multicolumn{2}{c}{Positive ranking score} & \multicolumn{2}{c}{Negative ranking score} \\
        \midrule
                    & \multicolumn{2}{c}{Stroke predictions, $D=1,897$} & \multicolumn{2}{c}{Non-stroke predictions, $D=342,133$} \\
        \midrule
                    & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ \\
        \midrule
        1.  & Ambulance & 1,680 & Tetanus & 4,378 \\        
        2.  & Blood clot & 895 & Pregnant & 8,749 \\        
        3.  & Left & 1,108 & Cut & 7,592 \\        
        4.  & Right & 1,050 & Bandage & 4,561 \\        
        5.  & Double vision & 84 & Amager (a location) & 23,776 \\        
        6.  & The words & 344 & O'clock & 94,436 \\        
        7.  & Suddenly & 783 & The emergency room & 42,809 \\        
        8.  & Arm & 709 & The police & 2,903 \\        
        9.  & Side & 1,139 & Swollen & 60,559 \\        
        10. & Stroke & 117 & Over the counter (OTC) & 4,641 \\        
        11. & Double & 113 & The neck & 30,151 \\        
        12. & Control & 134 & Fever & 112,586 \\        
        13. & Call & 39 & Prescription & 5,450 \\        
        14. & Numb & 94 & Centimetre & 12,026 \\        
        15. & Minutes & 763 & The knee & 8,875 \\        
        16. & Difficulties speaking & 44 & The pharmacy & 10,085 \\        
        17. & Haemorrhagic stroke & 133 & The stomach & 42,105 \\        
        18. & Hand & 297 & Psychiatric & 3,688 \\        
        19. & The ambulance & 521 & Pneumonia & 7,597 \\        
        20. & Slurred speech & 58 & Stomach pain & 10,551 \\        
        21. & Blood clots & 224 & Stool & 19,155 \\        
        22. & Fast & 663 & The ribs & 3,928 \\        
        23. & Express & 44 & Bleed & 10,501 \\        
        24. & Blood thinner & 259 & Bleeding & 24,313 \\        
        25. & Incoherent & 15 & Ribs & 2,941 \\        
        26. & Lopsided & 211 & Broken & 19,415 \\        
        27. & Reduced & 528 & Inflammation & 10,050 \\        
        28. & Hangs & 628 & Common cold & 8,127 \\        
        29. & Transient & 48 & Morning or morrow & 78,558 \\        
        30. & Not making sense & 14 & Swelling & 17,762 \\        
        \bottomrule
    \end{tabular}%
    }
\end{table}

@@

% }

>>>>

\chapter[a retrospective study on machine learning-assisted stroke recognition for medical helpline calls]{A Retrospective Study on Machine Learning-Assisted Stroke Recognition for Medical Helpline Calls}
\label{chp:paper-retrospective}
\ifthenelse{\equal{\skippapers}{true}}{}{

@@

\section{Introduction}
Stroke is a leading cause of disability and death worldwide \cite{cite1,cite2,cite3}. Effective treatment is time-sensitive, and an optimal outcome is more likely when treatment is administered within the first four and a half hours from stroke onset \cite{cite4,cite5}. The gateway to ambulance transport and hospital admittance is through prehospital telehealth services, including emergency medical call centres, nurse advice call lines, and out-of-hours health services. In the pre-hospital setting, the use of mobile stroke units has made it possible to deliver advanced treatment faster. As the mobile stroke unit is only dispatched to patients with a suspected stroke, the impact of mobile stroke units is directly influenced by accurate call-taker recognition of stroke \cite{cite6,cite7}. Call-takers who can rapidly and accurately recognize stroke are therefore crucial in facilitating prompt care in both pre-hospital and in-hospital settings.

Despite initiatives to improve stroke recognition \cite{cite8,cite9}, approximately half of all patients with stroke do not receive the correct triage for their condition from call-takers \cite{cite10,cite11,cite12}. Most initiatives aim to improve stroke recognition by call-takers via introducing more specific assessment tools \cite{cite8,cite9} or providing specialized training \cite{cite13}. Recent advances in machine learning technology might be applied to improve stroke recognition without requiring changes to the triaging approach, and machine learning aided identification of stroke has been suggested as a means of improving mobile stroke unit effectiveness \cite{cite7}. Real-time feedback from a machine learning model can improve the recognition of out-of-hospital cardiac arrest \cite{cite14,cite15}. Therefore, this study aimed to develop and assess the potential of machine learning in improving prehospital stroke recognition during medical helpline calls.

@@

This study has some limitations. First, the mapping of call recordings to electronic records was incomplete due to technical limitations in the CAD registry, which limited the number of calls available to us. Of note, there was no obvious pattern of bias related to the unmapped calls, and we included all calls with matching audio files, regardless of dispatcher performance. The results could potentially be improved if more calls were available for analysis. Second, calls without a call-taker indicated diagnostic category were not included in the validation and test data because the call-taker's performance could not be evaluated. Moreover, in exploratory analyses, the model performed worse on these calls, which might be attributed to differences in population characteristics (\cref{tab_retrospective:table3}). Finally, the ground truth stroke labelling relied on the patient-reported time of onset being exact; however, estimating the accuracy of the timestamps in DanStroke was impossible.

@@

In conclusion, using the largest collection of audio calls from patients with stroke to date, we developed a machine learning framework that significantly outperformed human call-takers in stroke recognition in medical helpline calls. The framework can assist human call-takers during medical helpline calls. Ideally, this would enable a higher recognition of patients with stroke in the prehospital setting, benefiting both patient outcomes and health service resource allocation.

@@

JW, JDH and LB share lead authorship and contributed equally to the original drafting of the paper, review and editing, data curation, formal analysis, methodology, validation, and conceptualisation. JDH and LB contributed to software development and funding acquisition. SNFB contributed to the conceptualisation, data curation, funding acquisition, supervision, and writing - review and editing. LM contributed to conceptualisation, funding acquisition, project administration, supervision, and writing - review and editing. MS contributed to supervision and writing - review and editing. HC contributed to the methodology, supervision, and writing - review and editing. CK contributed to the conceptualisation, funding acquisition, methodology, project administration, supervision, and writing - review and editing. HCC contributed to the conceptualisation, data curation, funding acquisition, methodology, project administration, supervision, and writing - review and editing. HCC and CK share last authorship. The three lead authors directly accessed and verified the underlying data reported in this manuscript. JW's affiliations are exclusively academic.

@@

% \todo[inline]{Insert Table 3}
\begin{table}[t]
    \centering
    \caption[Words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.]{English translation of words with the largest positive and negative ranking score in calls predicted as stroke and non-stroke, respectively.}
    \label{tab_retrospective:table3}
    \resizebox*{0.98\textwidth}{!}{%
    \begin{tabular}{l|lr|lr}
        \toprule
                    & \multicolumn{2}{c|}{Positive ranking score} & \multicolumn{2}{c}{Negative ranking score} \\
        \midrule
                    & \multicolumn{2}{c|}{Stroke predictions, $D=1,897$} & \multicolumn{2}{c}{Non-stroke predictions, $D=342,133$} \\
        \midrule
                    & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ & Word, $w$ \textit{(translated)} & Occurences, $D^{(w)}$ \\
        \midrule
        1.  & Ambulance & 1,680 & Tetanus & 4,378 \\        
        2.  & Blood clot & 895 & Pregnant & 8,749 \\        
        3.  & Left & 1,108 & Cut & 7,592 \\        
        4.  & Right & 1,050 & Bandage & 4,561 \\        
        5.  & Double vision & 84 & Amager (a location) & 23,776 \\        
        6.  & The words & 344 & O'clock & 94,436 \\        
        7.  & Suddenly & 783 & The emergency room & 42,809 \\        
        8.  & Arm & 709 & The police & 2,903 \\        
        9.  & Side & 1,139 & Swollen & 60,559 \\        
        10. & Stroke & 117 & Over the counter (OTC) & 4,641 \\        
        11. & Double & 113 & The neck & 30,151 \\        
        12. & Control & 134 & Fever & 112,586 \\        
        13. & Call & 39 & Prescription & 5,450 \\        
        14. & Numb & 94 & Centimetre & 12,026 \\        
        15. & Minutes & 763 & The knee & 8,875 \\        
        16. & Difficulties speaking & 44 & The pharmacy & 10,085 \\        
        17. & Haemorrhagic stroke & 133 & The stomach & 42,105 \\        
        18. & Hand & 297 & Psychiatric & 3,688 \\        
        19. & The ambulance & 521 & Pneumonia & 7,597 \\        
        20. & Slurred speech & 58 & Stomach pain & 10,551 \\        
        21. & Blood clots & 224 & Stool & 19,155 \\        
        22. & Fast & 663 & The ribs & 3,928 \\        
        23. & Express & 44 & Bleed & 10,501 \\        
        24. & Blood thinner & 259 & Bleeding & 24,313 \\        
        25. & Incoherent & 15 & Ribs & 2,941 \\        
        26. & Lopsided & 211 & Broken & 19,415 \\        
        27. & Reduced & 528 & Inflammation & 10,050 \\        
        28. & Hangs & 628 & Common cold & 8,127 \\        
        29. & Transient & 48 & Morning or morrow & 78,558 \\        
        30. & Not making sense & 14 & Swelling & 17,762 \\        
        \bottomrule
    \end{tabular}%
    }
\end{table}

@@

}
